---
title: "Using the fars package"
author: "Lucas Zamora"
date: "2020-07-04"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Package Description

`fars` package reads and plots Fatality Analysis Reporting System (FARS) data 
from the [National Highway Traffic Safety 
Administration](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).



## Functions

Five functions are available; all exported for the user:

Function                 | Description
-------------------------|-------------------------------------------
`make_filename()`        | Generate FARS filename for a given year
`fars_read()`            | Read a FARS csv file into a tibble
`fars_read_years()`      | Read FARS data for a vector of years (helper function used by `fars_summarize_years`)
`fars_summarize_years()` | Summarise FARS observation counts by year and month
`fars_map_state()`       | Draw a geographical map of accident sites for a specific US state and year


## Load Packages

```{r}
library(fars)
library(maps)
```



## Included data

Data files included within the package are:

```{r}
list.files(system.file("extdata", package = "fars"))
```


### Function `make_filename`

This function returns file name and path to the data files:

```{r}
yr <- 2015
make_filename(yr)
```

### Function `fars_read_years`

Ancillary function used by `fars_summarize_years`

```{r, eval=FALSE}
fars_read_years(2013)
```


### Function `fars_summarize_years`

This function summarizes yearly accidents data, by month.

```{r, eval=FALSE}
plot(fars_summarize_years(2015))
fars_summarize_years(c(2015, 2014))
```

### Function `fars_map_state`

The function displays a state map plot including accidents location
by year

```{r, eval=FALSE}
state_num <- 49 # Utah
yr <- 2014
fars_map_state(state_num, yr)
```
